```{r}
install.packages("dplyr")
```

```{r}
library(dplyr)
```

```{r}
cchs <- read.csv("cchs.csv")
```

```{r}
install.packages("ggplot2")
install.packages("GGally")

library(ggplot2)
library(GGally)
```

Analysis on perceived stress versus hours of sleep for people aged 18-24.

Todo:

1.  filter on age, on people who participated in survey
2.  Recategorize some variables like stress to "high risk" or "low risk", recategorize perceived mental health to "high risk" or "low risk", leaving the other variables as is
3.  Perform confounding analysis, add both confounders in the regression model

Filter data based on criteria:

DOSLP: people who participated in sleep survey

DHHGAGE: only people ages 18-24

```{r}
filtered_data <- cchs %>%
  filter(DOSLP != 2, DHHGAGE %in% c(3, 4), FSCDVAFS != 9)
         
```

```{r}
filtered_data <- filtered_data %>%
  select(GEN_015, SLPG005, GEN_020, INCDGPER, EHG2DVR3, GEN_005, FSCDVAFS)
```

```{r}
ggpairs(filtered_data)
```

```{r}
filtered_data <- filtered_data %>%
  filter(!GEN_015 %in% c(7, 8, 9))
```

```{r}
filtered_data <- filtered_data %>%
  filter(SLPG005 <= 10)
```

```{r}
filtered_data <- filtered_data %>%
  filter(!GEN_020 %in% c(7, 8))
```

```{r}
filtered_data <- filtered_data %>% 
  filter(INCDGPER <= 10)
```

```{r}
filtered_data <- filtered_data %>%
  filter(EHG2DVR3 <= 5)
```

```{r}
filtered_data <- filtered_data %>%
  filter(!is.na(GEN_005), !GEN_005 %in% c(7, 8))
```

```{r}
ggpairs(filtered_data)
```


Convert data to format desired:

GEN_015: 0 is low risk, 1 is high risk (1, 2 is excellent, very good)

GEN_020, 0 is low stress, 1 is high stress

GEN_005: 0 is low risk, 1 is high risk

```{r}
mutated_data <- filtered_data %>%
  mutate(
    GEN_015 = case_when(GEN_015 %in% c(1, 2) ~ 0, TRUE ~ 1),
    SLPG005 = case_when(SLPG005 < 6 ~ 0, SLPG005 %in% c(6, 7) ~ 1, SLPG005 > 7 ~ 2),
    GEN_020 = case_when(GEN_020 %in% c(1, 2, 3) ~ 0, GEN_020 %in% c(4, 5) ~ 1),
    GEN_005 = case_when(GEN_005 %in% c(1, 2) ~ 0, GEN_005 %in% c(3, 4, 5) ~ 1)
  )
```


Analysis:

```{r}
summary(filtered_data)

filtered_data %>%
  select_if(is.factor)
```

```{r}
contingency_table <- table(mutated_data$GEN_015, mutated_data$SLPG005)
print(contingency_table)

```

```{r}
chi_square_test <- chisq.test(mutated_data$GEN_015, mutated_data$SLPG005)
print(chi_square_test)
```

```{r}
contingency_table2 <- table(mutated_data$GEN_015, mutated_data$GEN_020)
```

```{r}
print(contingency_table2)
```

```{r}
chi_square_test2 <- chisq.test(mutated_data$GEN_015, mutated_data$GEN_020)
print(chi_square_test2)
```

```{r}
chi_square_test3 <- chisq.test(mutated_data$GEN_015, mutated_data$INCDGPER)
print(chi_square_test3)
```

```{r}
chi_square_test4 <- chisq.test(mutated_data$GEN_015, mutated_data$EHG2DVR3)
print(chi_square_test4)
```
```{r}
chi_square_test5 <- chisq.test(mutated_data$GEN_015, mutated_data$GEN_005)
print(chi_square_test5)
```
```{r}
chi_square_test6 <- chisq.test(mutated_data$GEN_015, mutated_data$FSCDVAFS)
print(chi_square_test6)
```

Multivariable analysis:

```{r}
library(stats)
```

```{r}
mutated_data$GEN_015 <- factor(mutated_data$GEN_015)
mutated_data$GEN_015 <- factor(mutated_data$GEN_005)
mutated_data$SLPG005 <- factor(mutated_data$SLPG005)
mutated_data$GEN_020 <- factor(mutated_data$GEN_020)
mutated_data$INCDGPER <- factor(mutated_data$INCDGPER)
mutated_data$EHG2DVR3 <- factor(mutated_data$EHG2DVR3)

mutated_data$FSCDVAFS <- factor(mutated_data$FSCDVAFS)
```

```{r}
model <- glm(GEN_015 ~ SLPG005 + GEN_020 + INCDGPER + EHG2DVR3 + FSCDVAFS, data = mutated_data, family = binomial())

```

```{r}
summary(model)
exp(coef(model))
```


```{r}
coefficients <- coef(model)
std_errors <- sqrt(diag(vcov(model)))

# Calculate the confidence intervals
z_value <- qnorm(1 - (0.05 / 2))  # for 95% CI; adjust accordingly for other levels
lower_bounds <- coefficients - z_value * std_errors
upper_bounds <- coefficients + z_value * std_errors

# Convert to odds ratios and their confidence intervals
or <- exp(coefficients)
ci_lower <- exp(lower_bounds)
ci_upper <- exp(upper_bounds)

# Combine into a matrix or data frame for easier viewing
ci_matrix <- cbind(or, ci_lower, ci_upper)
colnames(ci_matrix) <- c("Odds Ratio", "Lower CI", "Upper CI")

# Print the matrix
print(ci_matrix)
```

Logistical Regression result usually have:

odds ratio: predicting odds of having bad mental health

estimate: coefficient: how much log odds of the dependent variable are expected to increase or decrease with a one-unit increase in the predictor

std. error: standard error, smaller std error is more precise

z-value: determines whether the coefficient is significantly different from zero, if Z value is large, then predictor is significant

p-value: small indicates model is reliable.


```{r}
library(nnet)
```

Confounding analysis for GEN_005

```{r}
summary(multinom(SLPG005 ~ GEN_005, data = mutated_data))
```

```{r}
summary(glm(GEN_015 ~ GEN_005, data = mutated_data, family = binomial()))
```

Confounding analysis for GEN_020

```{r}
summary(glm(GEN_015 ~ GEN_020, data = mutated_data, family = binomial()))
```

```{r}
write.csv(mutated_data, './data.csv', row.names = FALSE)
```

